---
title: "Code suggestions for Jhovae"
output:
  pdf_document: default
  html_notebook: default
---

Below is an example of what we want to do with the Naiman data, but with some data from a different project. You'll just need to modify it a bit to get it to work with the Naiman data.
Also to run this you'll need to have the data file that is called below, and the soils data, both of which are on github.

We're going to try to do two things. First is called Nonmetric Multidimensional Scaling, which is a fancy way of saying Clustering. We'll calculate a distance matrix and then plot the ordination diagrams, which basically calculates & visualized clusters in the data, so we can see if data are more similar WITHIN a site than between sites.

Then we will do a diff type of ordination called a Correspondence Analysis which allows us to see how environmental parameters correlate with the clusters/spread in the species data across the sites.

Load packages:
```{r}
library(vegan)
library(MASS)
library(dplyr)
library(tidyr)
library(reshape2)
library(stringr)
```

You'll use the code you already have written up for Naiman data for this next step, but I'm including this here to get the example w these data to work:
```{r}
#load the data
mgp16<- read.csv("./data/2016_MGIM_ga_veg_subset.csv", 
                  header = TRUE)

sp.df<- dplyr::select(mgp16, site_numeral, Plot, Species, Cover)
sp.df$Species<- str_replace_all(sp.df$Species, " ", ".")

# convert to a sitexspecies matrix
site.sp.quad <- dcast(sp.df, site_numeral + Plot ~ Species, 
                     value.var ='Cover', FUN=mean)
site.sp.quad[is.na(site.sp.quad)] <- 0   #replace NA with 0
site.matrix <- site.sp.quad    # just renaming it so I don't have to change the 
#name that is called in everything below
```
You should have your own version of all of that above.

Here's the new stuff.

```{r}
# calc distance matrix
dis.bray <- vegdist(site.matrix,method="bray")

# id 'treatments' & replicates for use in later steps when need to have labels 
#for plotting
# here the "treatments" are just the two regions, but in Naiman it will be each
#different site.
treat=c(rep("Mongolia",66),rep("Inner Mongolia",38))  # create a vector of the site
#names, each repeated the number of times = number of data points for the site. (66+
#38 = 104)

# we'll also use this info in later steps.
sites<- site.sp.quad[1:104, 1]         # extract site IDs
plots<- site.sp.quad[1:104, 2]        # extract plot replicates
```

Use the distance matrix you created above to run the NMDS:
```{r}
# 
dis.mds <- metaMDS(dis.bray, trace = FALSE, autotransform = FALSE, shrink= FALSE)
dis.mds
```

Then once we've run that analysis, we need to extract some info from the object with the results so that we can use that
infor for plotting later.
```{r}
scores <- data.frame(scores(dis.mds, display = "sites")) # this extracts the scores
#you want to plot in ggplot

# next bind the scores to the site info that we created above
scores2<- cbind(sites, scores) 
```

Basic plotting of this info:   
```{r}
plot(dis.mds, type = "t")   # just plots the individual sampling plots 
# plots w a polygon drawn around the clusters
ordihull(dis.mds,groups=treat,draw="polygon",col="grey90",label=T)  
# labels all the sampling pltos (pretty messy and hard to read, so we prob don't want this)
ordihull(dis.mds,groups=sites,draw="polygon",col="grey90",label=T)   
```

Now we want to bring in the environmental data!
But we need to load it first.
Load the soils data file and calc basic summary stats for each site.
```{r}

```

