---
title: "Naiman codyn"
output: html_notebook
---
1. Data import --------
```{r}
############ 2008 transects all:
t09 <- read.csv("./data/transect09.csv", sep= ",", header = TRUE, stringsAsFactors = FALSE)
t09 <- t09 %>% dplyr::select( -Month, -Day)
# then need to create a summary of sp by transect:

t09sum <- t09 %>% count(SiteName, Transect, SpeciesCode)%>% mutate(Year = 2008)
t09sum <- t09sum[2:254,]
# remove NAs
t09sum <- na.omit(t09sum)
#restrict number of transects:
#t09sum <- t09sum %>% dplyr::filter(Transect <= 1)

  # change variable names to match 2018:
t09sum <- t09sum %>% mutate(Site = case_when(SiteName == "HeavyGrazing" ~ "Heavy",
                                             SiteName == "RedoControl" ~ "Control",
                                             SiteName == "DXG" ~ "DXG",
                                             SiteName == "Naiman" ~ "Naiman",
                                             SiteName == "Grazing" ~ "Grazing",
                                             TRUE ~ as.character(SiteName))) #else NA

########## 2018 transects :
nm18 <- read_excel("./data/2018.Data.Entry.xlsx", sheet = 1)
nm18 <- nm18 %>% dplyr::select(-Month, -Day, -X__1)
tbl_df(nm18)

t18sum <- nm18 %>% count(SiteName, Transect, SpeciesCode) %>% mutate(Year = 2018)
t18sum <- mutate(t18sum, Site = SiteName)

t18sum$SiteName<- as.factor(t18sum$SiteName)
#t18sum$Transect<- as.factor(t18sum$Transect)
t18sum$SpeciesCode<- as.factor(t18sum$SpeciesCode)
#t18sum$Site<- as.factor(t18sum$Site)
#
#Manually add the Naiman transect data for 2018:
nm08<- t09sum %>% filter(SiteName == "Naiman")

SpeciesCode = rep(c("AH" ,"CM", "LD", "CS", "SM", "SV", "LO", "SalixQ", "EE"), 2)
SiteName <- rep("Naiman", 18)
n <- as.numeric(rep(c(15, 40, 13, 6, 20, 30, 1, 5, 6),2))
Transect <- c(rep(1, 9), rep(2, 9))
Year <- as.numeric(rep(2018, 18))
Site<- as.character(rep("Naiman", 18))

cols<-cbind(SiteName, Transect, SpeciesCode, n, Year, Site)
dat<-as.data.frame(cols, stringsAsFactors = FALSE)
dat$n<-as.numeric(as.character(dat$n))
dat$Year<-as.numeric(as.character(dat$Year))
dat$Transect<- as.numeric(as.character(dat$Transect))
str(dat)



# combine the two years:
x <- bind_rows(dat, t18sum, t09sum)
x18 <- x %>% filter(Year == 2018)

# add unique ID as Replicate whichis Site&Transect combined:

x <- x %>% mutate(Replicate = paste(Site, Transect, sep = "."))


# remove Light grazing (and DXG?)

trt <- c("Grazing", "Heavy", "Control", "Naiman")
trt2 <- c("Grazing", "Heavy", "Control", "Naiman", "DXG") # to use DXG as reference

nmtrans <- x %>% dplyr::filter(Site  %in% trt)
nmdxg <- x %>% dplyr::filter(Site  %in% trt2)

# w just 3 transects each:
#nmveg3t<- nmtrans %>% filter(Transect <= 3)

# w just 3 transects each:
nmveg3t<- nmdxg %>% filter(Transect <= 3)


###### mg_imar transect data:
        
    # not using this right now
mgp16<- read.csv("./data/2016_MGIM_ga_veg_subset.csv", 
                 header = TRUE)

mgsum <- mgp16 %>% group_by(plat, site_original, site_numeral, Plot, Species) %>%
                   summarise(n = sum(Cover)) %>%
                   mutate(Replicate = paste(site_original, Plot, sep = "."))

t09.3t<- nmveg3t %>% filter(Year == 2008)


```

2. Community analysis by year:
# community structure: --------
```{r}
  #2008         ******* this isn't right now. the richness values are too high???
community_diversity(df= t09.3t, abundance.var = "n", replicate.var = "Site")
community_structure(df= t09.3t, abundance.var = "n", replicate.var = "Site")

  #2018
community_diversity(x18, abundance.var = "n", replicate.var = "SiteName")
community_structure(df= x18, abundance.var = "n", replicate.var = "SiteName")

  #mg_imar
community_diversity(df = mgsum, abundance.var = "n", replicate.var = "site_original")
community_structure(df= mgsum, abundance.var = "n", replicate.var = "site_original")

community_diversity(df = mgsum, abundance.var = "n", replicate.var = "plat")
community_structure(df= mgsum, abundance.var = "n", replicate.var = "plat")

```
# rank abundance curves: --------  NOT WORKING. ASK MEGAN.
```{r}

RAC_change(df = nmveg3t , 
           species.var = "SpeciesCode",
           time.var = "Year",
           abundance.var = "n",
           replicate.var = "Replicate")

gzd<- nmveg3t%>% filter(Site == "Grazing")
RAC_change(df = gzd , 
           species.var = "SpeciesCode",
           time.var = "Year",
           abundance.var = "n",
           replicate.var = "Replicate")




# RAC_difference(df= mgsum, 
#                species.var = "Species", 
#                abundance.var = "n", 
#                replicate.var = "Replicate", treatment.var = "plat")

RAC_difference(df = nmveg3t,
               species.var = "SpeciesCode",
               abundance.var = "n",
               replicate.var = "Replicate",
               time.var = "Year")

RAC_difference(df = x18,
               species.var = "SpeciesCode",
               abundance.var = "n",
               replicate.var = "Replicate", treatment.var = "SiteName")

```
# species abundance ----------------
```{r}
abundance_change(df = nmveg3t,
                 time.var = "Year",
                 species.var = "SpeciesCode",
                 abundance.var = "n",
                 replicate.var = "Replicate"
                 )
    
# abundance_difference(df = mgsum,
#                      species.var = "Species",
#                      abundance.var = "n",
#                      replicate.var = "Replicate", 
#                      treatment.var = "plat",
#                      pool = TRUE)

```

# ORDINATION 
  *2008*
# convert to a sitexspecies matrix
```{r}
# nmdxg or nmveg3t?
# sep by year?
nmdxg08<- nmdxg %>% filter(Year == "2008")
site.sp.quad <- cast(nmdxg08, Replicate ~ SpeciesCode, 
                     value='n', FUN=mean)
site.sp.quad <- as.data.frame(site.sp.quad) #convert from class "cast" to a df
site.sp.quad[is.na(site.sp.quad)] <- 0   #replace NA with 0
site.matrix <- as.matrix.cast_df(site.sp.quad) #convert to a matrix
```
# calc distance matrix
```{r}

dis.bray <- vegdist(site.matrix,method="bray")
# id 'treatments' = region
# treat=c(rep("Mongolia",66),rep("Inner Mongolia",38))
sites<- site.sp.quad[1:49, 1]
# plots<- site.sp.quad[1:104, 2]
```
# Run the MDS
```{r}

dis.mds <- metaMDS(dis.bray, trace = FALSE, autotransform = FALSE, shrink= FALSE)
# view the results:
dis.mds
# pull the scores from the results:
scores <- data.frame(scores(dis.mds, display = "sites")) # this extracts the scores for later plotting
treat=c(rep("Control",15),rep("DXG",3), rep("Grazing", 13), rep("Heavy", 6), rep("Naiman", 12))
scores2<- cbind(sites, scores, treat) # bind the scores to the site info
```
## NMDS Plot 

Plot of the two NMDS axes, symbols separated by country (shape), site (color) and plot replicate (shade):
to compare whether there wasovelarp in species composition 
```{r echo=FALSE, fig.width = 10, fig.height = 7}
ggplot(scores2, aes(x=NMDS1, y=NMDS2, color=factor(sites)))+
   geom_point(aes(shape= treat), size=5)+
   scale_fill_identity()+    # need a better color scale
   xlab("NMDS Axis 1")+
   ylab("NMDS Axis 2")
```

``` {r echo = FALSE, warning = FALSE}
plot(dis.mds, type = "t")
ordihull(dis.mds,groups=treat,draw="polygon",col="grey90",label=T)
```

  *2018*

# convert to a sitexspecies matrix
```{r}
# nmdxg or nmveg3t?
# sep by year?
nmdxg18<- nmdxg %>% filter(Year == "2018")
site.sp.quad <- cast(nmdxg18, Replicate ~ SpeciesCode, 
                     value='n', FUN=mean)
site.sp.quad <- as.data.frame(site.sp.quad) #convert from class "cast" to a df
site.sp.quad[is.na(site.sp.quad)] <- 0   #replace NA with 0
site.matrix <- as.matrix.cast_df(site.sp.quad) #convert to a matrix
```
# calc distance matrix
```{r}

dis.bray <- vegdist(site.matrix,method="bray")
# id 'treatments' = region
# treat=c(rep("Mongolia",66),rep("Inner Mongolia",38))
sites<- site.sp.quad[1:5, 1]
# plots<- site.sp.quad[1:104, 2]
```
# Run the MDS
```{r}

dis.mds <- metaMDS(dis.bray, trace = FALSE, autotransform = FALSE, shrink= FALSE)
# view the results:
dis.mds
# pull the scores from the results:
scores <- data.frame(scores(dis.mds, display = "sites")) # this extracts the scores for later plotting
treat=c("Control","Grazing", "Heavy", "Naiman", "Naiman")
scores2<- cbind(sites, scores, treat) # bind the scores to the site info
```
## NMDS Plot 

Plot of the two NMDS axes, symbols separated by country (shape), site (color) and plot replicate (shade):
to compare whether there wasovelarp in species composition 
```{r echo=FALSE, fig.width = 10, fig.height = 7}
ggplot(scores2, aes(x=NMDS1, y=NMDS2, color=factor(sites)))+
   geom_point(aes(shape= treat), size=5)+
   scale_fill_identity()+    # need a better color scale
   xlab("NMDS Axis 1")+
   ylab("NMDS Axis 2")
```

``` {r echo = FALSE, warning = FALSE}
plot(dis.mds, type = "t")
ordihull(dis.mds,groups=treat,draw="polygon",col="grey90",label=T)
```
